Dynamical properties of the Benguela Current as seen through a cold-spell algorithm.

### 2019-03-20 Write_up/Writing/dissertation.Rmd
Geostrophic_currents

Reliable datasets are often difficult to maintain and obtain. In the field of biological science research, it is often abiotic data products that are impaired thus leading to the use of satellite derived Sst products for the study of specific fields such as coastal and marine ecology as well as oceanography. These datasets however, may be inaccurate along the coast and local in situ collected data should be used when possible (Smit et al., 2013). Remotely sensed data products are obtained by several different organisations worldwide. Satellites launched into space detect the infra-red light emitted by the ocean which is then taken as a proxy from SST. Given the extensive distance at which these satellites record data, a certain level of coarseness is displayed within the dataset. This is demonstrated as the size of the pixels one sees in the final product.

@@

The Benguela system is bounded by the equatorward and poleward ends by warm water regimes. These currents are shown to be forced by local wind stress fields off southwest Africa[@nelson ; @fennel]. This EBUS streches along the south-western coast of Africa from southern Angola to Cape Agulhas [@cole]. Water is coldest and most persistent off the Namibian coast at 25°S and weakest at 17°S [@Lutjeharms].  The upwelling period in the benguela north of 25°C  persist throughout the entire year but peaks during August. The Benguela upwelling region is split into northern and southern systems by a zone of intense perennial upwelling activity in the Luderitz region within the Namibian region [@shannon; @cole]. Several research studies have shown extreme warming and cooling periods occurring in the northern and southern benguela systems [@stander; @walker; @shannon1984; @boyd; @brainard]. Research studies done by Walter (1937) also show anomously warm periods which may have resulted as a lack of upwelling or poor upwelling events. During the periods of warming the equatorward wind stress over the northern region of the benguela system was stronger than usual; yet upwelling was suppressed by the southward movement of warm saline waters from the Angolan region. In the southern benguela region, continental shelf bathymetry and strong upwelling favourable winds are known to result in more intense upwelling events. Upwelling occuring along the south coastline of the of Southern Africa are particularly generated by local winds.

>>>>

Reliable datasets are often difficult to maintain and obtain. In the field of biological science research, it is often abiotic data products that are impaired thus leading to the use of satellite derived SST products for the study of specific fields such as coastal and marine ecology as well as oceanography. These datasets however, may be inaccurate along the coast and local in situ collected data should be used when possible (Smit et al., 2013). Remotely sensed data products are obtained by several different organisations worldwide. Satellites launched into space detect the infra-red light emitted by the ocean which is then taken as a proxy from SST. Given the extensive distance at which these satellites record data, a certain level of coarseness is displayed within the dataset. This is demonstrated as the size of the pixels one sees in the final product.

@@

The Benguela system is bounded by the equatorward and poleward ends by warm water regimes. These currents are shown to be forced by local wind stress fields off southwest Africa[@nelson; @fennel]. This EBUS streches along the south-western coast of Africa from southern Angola to Cape Agulhas [@cole]. Water is coldest and most persistent off the Namibian coast at 25°S and weakest at 17°S [@Lutjeharms].  The upwelling period in the benguela north of 25°C  persist throughout the entire year but peaks during August. The Benguela upwelling region is split into northern and southern systems by a zone of intense perennial upwelling activity in the Luderitz region within the Namibian region [@shannon; @cole]. Several research studies have shown extreme warming and cooling periods occurring in the northern and southern benguela systems [@stander; @walker; @shannon1984; @boyd; @brainard]. Research studies done by Walter (1937) also show anomously warm periods which may have resulted as a lack of upwelling or poor upwelling events. During the periods of warming the equatorward wind stress over the northern region of the benguela system was stronger than usual; yet upwelling was suppressed by the southward movement of warm saline waters from the Angolan region. In the southern benguela region, continental shelf bathymetry and strong upwelling favourable winds are known to result in more intense upwelling events. Upwelling occuring along the south coastline of the of Southern Africa are particularly generated by local winds.
